Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-02-07 response-to-reviews.Rmd

We are pleased to submit a revised version of our manuscript, "Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications" (NPH-TR-2021-38020), for consideration for publication in *New Phytologist*. We have addressed all points raised by the reviewers, as detailed below. The most significant changes include:

- ...

- ...

Regarding production of the figures, below are some specific criteria:
<!--To ensure rapid handling of your manuscript we would be grateful if you could confirm, in the cover letter to the Editor, whether or not there are any important elements of the figures that need to be retained (eg. colours used, plant species drawn, layout and spacing, symbols, etc) and if you have specific instructions that you would like the illustrator to take on board.-->

- It will generally be fine to adjust colors, layout, and spacing on the figures.
- We really like Figure 1 and would like to retain the background painting (by the lead author)
- On Fig. 4, please adjust colors on panel (a) to correspond to those in panel (b), where "min MIN" color should match "understory" and "max DBH" should match "canopy". The same color scheme can be applied to all panels in (a), and there is no need to keep the line type differences (both can be solid).

### 2022-02-07 vertical-thermal-review.Rmd
fix Piponiot reference
@NidhiVinod , I've synced the updated reference in Zotero.

Canopy trees account for the majority of forest ecosystem water and carbon cycling.
While studies partitioning transpiration across forest vertical gradients are rare, both evapo-transpiration (ET) and transpiration have been shown to increase with height in a *Picea abies* forest in Germany, such that the upper half of the canopy contributed an estimated 80% of daytime ET [@staudtVerticalStructureEvapotranspiration2011]. 
Similarly, in a tropical forest in the Brazilian Amazon, canopy and subcanopy trees jointly contributed approximately 93% of ET, or 88% of transpiration [Fig `r fig_ecosystem`a, @kunertRevisedHydrologicalModel2017], and trees >33 cm DBH contributed >80% of transpiration in a forest in Costa Rica [@mooreUpscalingTranspirationDiverse2018]. 
In terms of C cycling, it has been estimated that canopy strata contribute $\ge$ 64% of net daytime CO~2~ uptake [i.e., GPP - ecosystem respiration, including from soil, @missonPartitioningForestCarbon2007]. 
Large trees also dominate in terms of woody above ground net primary productivity ($ANPP_{woody}$) and mortality ($M_{woody}$), with trees $\geq$ 10cm DBH usually contributing > 85% of $ANPP_{woody}$ and $M_{woody}$ across 25 globally distributed forests [Fig. `r fig_ecosystem`b, @piponiotSizerelatedTreeBiomassinreview in review].

![**Figure `r fig_ecosystem`. Vertical partitioning of (a) evapotranspiration and (b) C fluxes in tropical forests.** Panel (a) is from @kunertRevisedHydrologicalModel2017; panel (b) presents averages for 14 tropical forests from @piponiotSizerelatedTreeBiomassinreview in review.](schematics/ecosystem vertical partitioning/ecosystem partitioning.png)

>>>>

Canopy trees account for the majority of forest ecosystem water and carbon cycling.
While studies partitioning transpiration across forest vertical gradients are rare, both evapo-transpiration (ET) and transpiration have been shown to increase with height in a *Picea abies* forest in Germany, such that the upper half of the canopy contributed an estimated 80% of daytime ET [@staudtVerticalStructureEvapotranspiration2011]. 
Similarly, in a tropical forest in the Brazilian Amazon, canopy and subcanopy trees jointly contributed approximately 93% of ET, or 88% of transpiration [Fig `r fig_ecosystem`a, @kunertRevisedHydrologicalModel2017], and trees >33 cm DBH contributed >80% of transpiration in a forest in Costa Rica [@mooreUpscalingTranspirationDiverse2018]. 
In terms of C cycling, it has been estimated that canopy strata contribute $\ge$ 64% of net daytime CO~2~ uptake [i.e., GPP - ecosystem respiration, including from soil, @missonPartitioningForestCarbon2007]. 
Large trees also dominate in terms of woody above ground net primary productivity ($ANPP_{woody}$) and mortality ($M_{woody}$), with trees $\geq$ 10cm DBH usually contributing > 85% of $ANPP_{woody}$ and $M_{woody}$ across 25 globally distributed forests [Fig. `r fig_ecosystem`b, @piponiotSizerelatedTreeBiomass2022].

![**Figure `r fig_ecosystem`. Vertical partitioning of (a) evapotranspiration and (b) C fluxes in tropical forests.** Panel (a) is from @kunertRevisedHydrologicalModel2017; panel (b) presents averages for 14 tropical forests from @piponiotSizerelatedTreeBiomass2022.](schematics/ecosystem vertical partitioning/ecosystem partitioning.png)
